Pool Model Paper

### 2023-12-06 pool_model.tex
Corrections: mutual inhibition od species

The bacterial growth curves usually consists of different consecutive phases~\cite{buchanan_when_1997}.

@@

In this state the cultures accumulate waste products of metabolitic activity that leads to cell death of the bacteria.

@@

In the system where all needed nutritients are in excess except for one the total growth is linear to the concentration of this liminting nutritient~\cite{hibbing_bacterial_2010, monod_growth_1949}.

@@

        The red curve describes the growth of the bacteria $A$ $N_A = \sum_{k} x_A^k = L_A+G_A$ with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
        The blue curve describes the growth of the bacteria $B$ $N_B = \sum_{k} x_B^k = L_B+G_B$ with model parameters $\lambda_A=5\cdot 10^{-2}$, $\alpha_A=2.5$.

@@

\subsubsection{Species Inhibition}
\cmt{\cite{lenski_coexistence_1986}: Modeling of the 2 sp. system with 1resource and 1 inhibitor.}
The simplest way to introduce inhibition of species A by species B is via a growth rate of species A depending on the abundance of species B. 

@@

It is straight forward to make the mechanism of inhibition more explicit by, e.g., introducing a waste pool. 

@@

This corresponds to the case $W\sim G$, where $W$ is the waste (or whatever one chooses to call it) of species A affecting species B. 
If the waste accumulates one needs to replace $G$ by $\int G dt$. Keeping it simple yields:

@@

We can separate three cases: 

@@

    \includegraphics[width=0.6\textwidth]{Figures/pool_model_1pool_inhib.pdf}

@@

        The two species system where species inhibit each other defined with equations~(\ref{eq:2sp_1pool_inhibit_unitless1}-\ref{eq:2sp_1pool_inhibit_unitless2}).

>>>>

The bacterial growth curves usually consist of different consecutive phases~\cite{buchanan_when_1997}.

@@

In this state the cultures accumulate waste products of metabolic activity that leads to cell death of the bacteria.

@@

According to the works of Jacques Monod, in the system where all needed nutritients are in excess except for one the total growth is linear to the concentration of this liminting nutritient~\cite{monod_growth_1949}.

@@

        The blue curve describes the growth of the bacteria $B$ ($N_B = \sum_{k} x_B^k = L_B+G_B$) with model parameters $\lambda_A=5\cdot 10^{-2}$, $\alpha_A=2.5$.

@@

\subsubsection{Mutual Inhibition}
Another special case of the species competition is when the produced toxin does not cause the death of the bacteria but rather inhibit its growth.
Such situation with 2 bacteria species, limiting resource and produced inhibitor was modeled, e.g., by Lenski and Hattingh~\cite{lenski_coexistence_1986}.
In our case, we would like to look at rather simplified case reducing the number of variables of the system.
The easiest way to introduce inhibition of species $A$ by species $B$ is via a growth rate of species $A$ depending on the abundance of species $B$.

@@

It is straight forward to make the mechanism of inhibition more explicit by, e.g., introducing a toxin pool. 

@@

This corresponds to the case $T\sim G$, where $T$ is the toxin produced. 
If the toxin accumulates one needs to replace $G$ by $\int G dt$. Keeping it simple yields:

@@

We can separate three cases (see Figure~\ref{fig:1pool_2sp_inhibit}): 

@@

    \includegraphics[width=0.5\textwidth]{Figures/pool_model_1pool_inhib.pdf}

@@

        The two-species system mutually inhibiting each other is defined with equations~(\ref{eq:2sp_1pool_inhibit_unitless1}-\ref{eq:2sp_1pool_inhibit_unitless2}).
